### 2017-08-31 paper.tex
:art: Water entrainment

\subsection{Initiation phase}
Comparing the initial evolution of runout between the loose and the dense cases ($t = 0$ to $3\tau_c$), see~\cref{fig:slope_loose_5,fig:dense-s5-r09}, the loose granular columns evolve faster than their dense counterpart. When a granular material is sheared in the submerged conditions, it generates negative pore water pressure initially due to the undrained conditions. In the undrained conditions, the dilation movement (volume expansion) of the granular assembly by shearing self-generates negative excess pore pressure inside the pores. The fluid inside the pores does not have time to seep in or out from the free water outside the column by the internally generated pressure gradient between the pore space and the outside free water. The negative excess pore pressures produce larger interparticle forces by bringing the grains together. Macroscopically this results in increase in the mean effective stress and provides large shear resistance temporarily until the negative excess pore pressure dissipates with time. Large negative pore pressures were observed for the dense cases. The length of the failure plane increases with the slope angle, thereby increasing the volume of body with negative pore pressure that needs to be dissipated before the column starts to move. Overcoming the region of large negative pore pressure translates to smaller kinetic energy and eventually a shorter runout distance for the dense granular columns.

@@

>>>>

\subsection{Initiation phase}
Comparing the initial evolution of runout between the loose and the dense cases ($t = 0$ to $3\tau_c$), see~\cref{fig:slope_loose_5,fig:dense-s5-r09}, the loose granular columns evolve faster than their dense counterpart. When a granular material is sheared in the submerged conditions, it generates negative pore water pressure initially due to the undrained conditions. In the undrained conditions, the dilation movement (volume expansion) of the granular assembly by shearing self-generates negative excess pore pressure inside the pores. The fluid inside the pores does not have time to seep in or out from the free water outside the column by the internally generated pressure gradient between the pore space and the outside free water. The negative excess pore pressures produce larger interparticle forces by bringing the grains together. Macroscopically this results in increase in the mean effective stress and provides large shear resistance temporarily until the negative excess pore pressure dissipates with time. Large negative pore pressures were observed for the dense cases. Overcoming the region of large negative pore pressure translates to smaller kinetic energy and eventually a shorter runout distance for the dense granular columns.

@@

\subsection{Water entrainment}
Comparing the snapshots of the granular flow at time $t=3\tau_c$ for the dense column cases~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5,fig:dense-s5-r09}, it can be seen that the overall shapes of the flow front are different.~\Cref{fig:slope-outline} shows the boundary of the flowing mass and its evolution with time for the slope angle \si{5}{\degree} case. It shows that the dense granular columns exhibit a more acute angle of attack (flow front) in contrast to a more parabolic profile for the loose granular columns. This difference in angle of flow front changes the ratio of amount of water entering the flowing granular mass to the water interacting with the surface of the granular mass.~\Cref{fig:water_entrainment_loose_dense} shows the horizontal velocity of the fluid and streamlines at the flow front of the dense and loose granular columns for the slope angle of \si{5}{\degree}. The density of streamlines entering the granular front is significantly higher for the loose column case, in comparison to the dense column case. This clearly shows the effect of the shape of flow front on water entrainment. More water entrainment results in increase in pore space (or decrease in solid fraction) at the flow front, which in turn reduces the interparticle shear resistance.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.9\textwidth]{figs/slope5_dense_loose_outline}
\caption{Profile outline evolution for dense (solid line) and loose (dashed line) cases for a slope of \si{5}{\degree}}
\label{fig:slope-outline}
\end{figure}

%
\begin{figure}[!ht]
	\centering
\begin{subfigure}[b]{0.85\textwidth}
	\centering
    \includegraphics[width=\textwidth]{figs/slope5_water_entrainment/slope5_dense_15d_water_entrainment}
    \caption{streamlines at the flow front dense case (contours show velocity of fluid entrainment at the flow front)}
    \label{fig:water_entrainment_a08_dense}
\end{subfigure}\\
\begin{subfigure}[b]{0.85\textwidth}
	\centering
    \includegraphics[width=\textwidth]{figs/slope5_water_entrainment/slope5_loose_15d_water_entrainment}
    \caption{streamlines at the flow front loose case (contours show velocity of fluid entrainment at the flow front)}
    \label{fig:water_entrainment_a08_loose}
\end{subfigure}
\caption{Water entrainment for a slope of \si{5}{\degree} (dense and loose cases).}
\label{fig:water_entrainment_loose_dense}
\end{figure}
